---
title: "DATA 607 Project 3 Analysis"
author: "Susanna Wong"
date: "2023-03-15"
output:
  html_document:
    theme: cerulean
    code_folding: hide
    toc: true
    toc_float: true
editor_options: 
  markdown: 
    wrap: sentence
---

# Introduction
The goal of the project is to use data to answer the question, "Which are the most valued data science skills?".
We web scrap a major job search engine (ex: Indeed, Linkedin, Glassdoor, etc) for job postings.
Create a large csv file that contains the job title, job URL, company name, job salary, and job description.
Then, we will create a list of technical skills, and soft skills, and find the top 5 most frequent technical skills and soft skills on the job postings.
The most valued data science skills are the skills that appear most often on job postings.

# Load the CSV file

We webscrap job postings from LinkedIn using Octoparse. 
```{r}
library(stringr)
tidied_Linkedin <- read.csv('https://raw.githubusercontent.com/suswong/DATA-607-Project-3/main/tidied_Linkedin_version4.csv')
#colnames(tidied_Linkedin)
```

# Search {.tabset}

## Techincal Skills
```{r}
technical_skills <- tidied_Linkedin[c('Job_title','Company','Job_location_City','Job_location_State','Seniority_level','Employment_type','Job_function','Industries','Job_description')]

technical <- c("(?i)python", "(?i)java", "(?i)javascript", "(?i)sql", "(?i)\\bR\\b", "(?i)spss","(?i)stata", "(?i)sas","(?i)scala","(?i)matlab", "(?i)\\bC\\b","(?i)\\bC++\\b", "(?i)Swift", "(?i)Julia","(?i)Hadoop","(?i)Spark","(?i)Pig","(?i)Tableau","(?i)REDCap", "(?i)Qualtrics", "(?i)Power BI", "(?i)Dedoose", "(?i)Atlas TI", "(?i)NVivo", "(?i)MPlus", "(?i)Mixor", "(?i)dbt", "(?i)BigQuery|big query", "(?i)Superset", "(?i)Baseten", "(?i)Airflow","(?i)ETL/ELT pipelines","(?i)Google Analytics", "(?i)Parse.ly", "(?i)Chartbeat","(?i)Excel", "(?i)Plotly", "(?i)Google Data Studio", "(?i)Looker","(?i)Spotfire","(?i)Smartsheet","(?i)\\bBI\\b","(?i)C#|C#.net","(?i)Statistical analysis","(?i)data analysis","(?i)Machine Learning","(?i)data visual","(?i)wrangling","(?i)mathematics","(?i)programming","(?i)statistics","(?i)big data","(?i)Artificial intelligence","(?i)Deep learning")

for(i in technical){
   technical_skills[i] <- +str_detect(apply(technical_skills, 1, paste0, collapse = " ", ignore_case = TRUE), technical[which(technical == i)])
}
#tidied_Linkedin


# technical_skills <- tidied_Linkedin[c('Job_title','Job_description')]
# technical_skills$Python    <- grepl("Python", technical_skills$Job_description, ignore.case=TRUE)
# technical_skills$sql    <- grepl("sql", technical_skills$Job_description, ignore.case=TRUE)
#technical_skills$R    <- grepl("\\bR\\b", technical_skills$Job_description, ignore.case=TRUE)
#sum(technical_skills$R)

```

## Programming Skills
```{r}
programming_skills <- tidied_Linkedin[c('Job_title','Company','Job_location_City','Job_location_State','Seniority_level','Employment_type','Job_function','Industries','Job_description')]

programming <- c("(?i)python", "(?i)java", "(?i)javascript", "(?i)sql", "(?i)\\bR\\b", "(?i)spss","(?i)stata", "(?i)sas","(?i)scala","(?i)matlab", "(?i)\\bC\\b","(?i)\\bC++\\b", "(?i)Swift", "(?i)Julia","(?i)Hadoop","(?i)Spark","(?i)Pig","Go", "(?i)Kotlin","(?i)PHP","(?i)C#|C#.net","(?i)Ruby","(?i)TypeScript","(?i)HTML","(?i)CSS","(?i)NOSQL","(?i)Rust","(?i)Perl")

for(i in programming){
   programming_skills[i] <- +str_detect(apply(programming_skills, 1, paste0, collapse = " "), programming[which(programming == i)])
}

# Source: https://www.simplilearn.com/best-programming-languages-start-learning-today-article

```

### Non-Technical Skills

```{r}
nontechnical_skills <- tidied_Linkedin[c('Job_title','Company','Job_location_City','Job_location_State','Seniority_level','Employment_type','Job_function','Industries','Job_description')]

nontechnical <- c("(?i)communication", "(?i)critical thinking", "(?i)problem solving", "(?i)business acumen", "(?i)storytell", "(?i)adaptability","(?i)team player", "(?i)product understanding","(?i)innovation","(?i)collaboration", "(?i)technical")

for(i in nontechnical){
   nontechnical_skills[i] <- +str_detect(apply(nontechnical_skills, 1, paste0, collapse = " "), nontechnical[which(nontechnical == i)])
}

# Source: https://www.simplilearn.com/best-programming-languages-start-learning-today-article

```

# Frequent Skills {.tabset}

## Technical Skills
```{r}
library(tidyverse)
long_table <- technical_skills %>% 
  pivot_longer(cols = !c('Job_title','Company','Job_location_City','Job_location_State','Seniority_level','Employment_type','Job_function','Industries','Job_description'), names_to = "skill", values_to = "count" )

frequent_skill <- long_table %>% 
  filter(count == 1)
frequent_skill <- frequent_skill  %>% 
  count(skill) %>% 
  mutate(percentage = (n / nrow(tidied_Linkedin))*100)

library(DT)
datatable(frequent_skill)
```

## Programming

Based on 
```{r}
programming_long_table <- programming_skills %>% 
  pivot_longer(cols = !c('Job_title','Company','Job_location_City','Job_location_State','Seniority_level','Employment_type','Job_function','Industries','Job_description'), names_to = "skill", values_to = "count" )

frequent_programming_skill <- programming_long_table %>% 
  filter(count == 1)
frequent_programming_skill <- frequent_programming_skill  %>% 
  count(skill) %>% 
  mutate(percentage = (n / nrow(tidied_Linkedin))*100)

datatable(frequent_programming_skill)
```
## Non-technical 
```{r}
nontechnical_long_table <- nontechnical_skills %>% 
  pivot_longer(cols = !c('Job_title','Company','Job_location_City','Job_location_State','Seniority_level','Employment_type','Job_function','Industries','Job_description'), names_to = "skill", values_to = "count" )

frequent_nontechnical_skill <- nontechnical_long_table %>% 
  filter(count == 1)
frequent_nontechnical_skill <- frequent_nontechnical_skill  %>% 
  count(skill) %>% 
  mutate(percentage = (n / nrow(tidied_Linkedin))*100)

datatable(frequent_nontechnical_skill)
```

# Data Visualization {.tabset}

## Technical Skills
```{r}
ggplot(frequent_skill , aes(x = reorder(skill, percentage), y = percentage)) + 
  geom_bar(stat = 'identity', skill= "blue" , fill = 'lightblue') + 
  coord_flip() + ggtitle("Skills Used Most Often by Percentage") + 
  ylab("Percentage of skill Used") + 
  xlab("skill") 
```
## Programming 
```{r}
```
Sources

https://stackoverflow.com/questions/73640981/using-grep-on-multiple-columns-to-create-new-variable-in-r

https://stackoverflow.com/questions/44530029/how-to-ignore-case-when-using-str-detect